---
title: "Genes of interest"
output: html_notebook
---

```{r}
library("tidyverse")
library("here")
library("DESeq2")
```

```{r}
res_df <- as.data.frame(res)

# # baseMean: average of the normalized count taken over all samples
# log2FoldChange: fold change of genes in the treated condition when compared with the control (positive = upregulated, negative = downregulated)
## all fold changes are results for the "treated" because the treated is compared to the control
# lfcSE: standard error estimates of the log2FoldChange
# stat: 
# pvalue: 
# padj: corrected p value for multiple testing
```

```{r}
# Which genes are the most upregulated after treatment (in comparison to before)?
res_df %>%
  arrange(desc(log2FoldChange))

res_gene_ids <- rownames_to_column(res_df, var = "gene_id")

res_gene_ids %>%
  select(gene_id, log2FoldChange) %>%
  filter(log2FoldChange > 9.5) %>%
  arrange(desc(log2FoldChange)) %>%
  mutate(gene_id = case_when(
    gene_id == "MSTRG.10596|ENSG00000214425|ENSG00000266504|ENSG00000267246|ENSG00000285668|ENSG00000291175|LRRC37A4P" ~ "LRRC37A4P",
    gene_id == "ENSG00000102245|CD40LG" ~ "CD40L",
    gene_id == "ENSG00000162654|GBP4" ~ "GBP4",
    gene_id == "ENSG00000271503|CCL5" ~ "CCL5",
    gene_id == "ENSG00000164400|CSF2" ~ "CSF2",
    gene_id == "ENSG00000130477|UNC13A" ~ "UNC13A",
    gene_id == "MSTRG.12003|ENSG00000105371|ICAM4" ~ "ICAM4",
    gene_id == "ENSG00000163121|NEURL3" ~ "NEURL3",
    gene_id == "MSTRG.5668|ENSG00000139323|ENSG00000257594|ENSG00000259075|POC1B-GALNT4" ~ "POC1B-GALNT4",
    gene_id == "MSTRG.15256|ENSG00000275708|ENSG00000277437|MIR3648-1" ~ "MIR3648-1"
  )) %>%
  ggplot() + aes(x = gene_id, y = log2FoldChange) + geom_col()
```
*1. ENSG00000214425|ENSG00000266504|ENSG00000267246|ENSG00000285668|ENSG00000291175|LRRC37A4P*
- leucine rich repeat containing 37 member A4, pseudogene
- pseudogenes are prevented from encoding mRNA and thus non-protein coding -> so, does that make it irrelevant for us?

https://www.ensembl.org/Homo_sapiens/Gene/Summary?g=ENSG00000214425;r=17:45508109-45550335;t=ENST00000581296
https://www.genome.gov/genetics-glossary/Pseudogene


*2. ENSG00000102245|CD40L*
- expected as this is the ligand we were expressing

https://www.proteinatlas.org/ENSG00000102245-CD40LG


*3. ENSG00000162654|GBP4*
- Guanylate binding protein 4 (aka Mpa2)

Protein function
- binds GTP, GDP and GMP (ligands)
- hydrolizes GTP very efficiently
- plays a role in erythroid differentiation

Molecular function: hydrolase

Gene summary: Guanylate-binding proteins, such as GBP4, are induced by interferon and hydrolyze GTP to both GDP and GMP.

Tissue RNA expression: immune cells
- Single cell type specificity: cell type enhanced (Theca cells, adipocytes, endothelial cells, monocytes)
- Single cell type expression cluster (endothelial cells - angiogenesis (mainly))
- Cell type enriched (breast and kidney - endothelial cells)

Cancer and cell lines
- prognostic marker in renal cancer (unfavorable) and ovarian cancer (favorable)
- cell line expression cluster: non-specific - antiviral immune response (mainly)

https://www.proteinatlas.org/ENSG00000162654-GBP4


*4. ENSG00000271503|CCL5*
- C-C motif chemokine ligand 5
- high expression in a subset of immune cells 

Tissue RNA expression
- tissue expression cluster: immune cells - immune response (mainly)
- single cell type specificity: NK- and T- cells
- single cell type expression cluster: NK-cells - immune response regulation (mainly)

Immune cells
- immune cell specificity: memory CD8 T-cell, gdT-cell
- immune cell expression cluster: GdT cells - mainly adaptive immune response

Cancer and cell lines:
- prognostic marker in renal cancer (unfavorable), endometrial cancer (favorable), and melanoma (favorable)


*5. ENSG00000164400|CSF2*

*6. ENSG00000130477|UNC13A*

*7. ENSG00000105371|ICAM4*

*8. ENSG00000163121|NEURL3*

*9. ENSG00000139323|ENSG00000257594|ENSG00000259075|POC1B-GALNT4*

*10. ENSG00000275708|ENSG00000277437|MIR3648-1*



```{r}
# Which genes are the most downregulated after treatment (in comparison to before)?
res_df %>%
  arrange(log2FoldChange)
```
1. ENSG00000202198|ENSG00000283293|7SK

2. ENSG00000129226|ENSG00000161956|ENSG00000161960|ENSG00000209582|ENSG00000238917|ENSG00000264772|ENSG00000277957|ENSG00000277985|SENP3-EIF4A1

Long Non-coding RNA SENP3-EIF4A1 Functions as a Sponge of miR-195-5p to Drive Triple-Negative Breast Cancer Progress by Overexpressing CCNE1 (https://www.frontiersin.org/articles/10.3389/fcell.2021.647527/full)

3. ENSG00000067900 | ROCK1
- negative regulator of VEGF-induced angiogenic endothelial cell activation

https://www.proteinatlas.org/ENSG00000067900-ROCK1

4. ENSG00000085719 | CPNE3
- Calcium-dependent phospholipid-binding protein that plays a role in ERBB2-mediated tumor cell migration in response to growth factor heregulin stimulation

https://www.proteinatlas.org/ENSG00000085719-CPNE3

5. ENSG00000150712|ENSG00000199731




